---
title: "Figure1"
author: "Prerana Parthasarathy"
date: "2023-06-12"
output: html_document
---

```{r setup, include=FALSE}
library(survminer)
library(survival)
library(dplyr)
set.seed(7)
```

## Figure 1E: Change in SNV count for each Pt colored by response (36 paired pts)
###           input: mutation counts file 


```{r input}
data_dir<-"Checkmat153_Immunogenicity_NatureMedicine/data/"
mut_counts_file<-"MUTATIONS_COUNT_NEOANTIGENS.txt"
mut_counts<-read.delim(paste0(data_dir,mut_counts_file),sep="\t",header=TRUE)
#out_dir<-"Checkmat153_Immunogenicity_NatureMedicine/data/plots/Figure1/"
```

###           input: clinical response data
```{r input2}
clinical_data<-read.delim(paste0(data_dir,"bms153dat-021918.csv"),sep=",",header=TRUE)
clinical_data_response<-unique(clinical_data[,c("PatientID","myBOR","OS_event","OS_time_Mon","PFS_event","PFS_time")])
clinical_data_response$myBOR[which(is.na(clinical_data_response$myBOR))]<-"PD"
clinical_data_response$color<-NA
clinical_data_response$color[which(clinical_data_response$myBOR == "CR/PR")]<-"green"
clinical_data_response$color[which(clinical_data_response$myBOR == "SD")]<-"orange"
clinical_data_response$color[which(clinical_data_response$myBOR == "PD")]<-"red"

```

## Including Plots

### Figure 1E - Good

```{r SNV_change_in_muts_1E, echo=FALSE}
mut_counts<-mut_counts[order(mut_counts$SNV_change_in_mutations),]
mut_counts_response<-merge(mut_counts,clinical_data_response,by.x="patient",by.y="PatientID",all.x=TRUE)
mut_counts_response<-mut_counts_response[order(mut_counts_response$SNV_change_in_mutations),]


#pdf(paste0(out_dir,"fig1E_SNV_change_in_mutations.pdf"),height=7,width=10)
barplot(mut_counts_response$SNV_change_in_mutations, col=mut_counts_response$color, border=mut_counts_response$color, space=0.3,
main = "Reduced SNVs is associated with clinical response", ylab="Change in Mutations (SNV)",names=mut_counts_response$patient,las=2,
cex.axis=1.2, cex.lab=1.7,legend.text= c( "PR: Partial Response", "SD: Stable Disease", "PD: Progressive Disease"),
        args.legend=list(title="Best Overall Response", fill=c("green","orange",  "red"),x="topright",inset = c(0, 0.5),cex=1.5))
#dev.off()
```

### Figure 1F and 1G
survival OS and PFS of SNV_Change_in_mutations - split by median. 

```{r OS for groups, echo=FALSE}
mut_counts_response$groups<-NA
v<-median(mut_counts_response$SNV_change_in_mutations)
mut_counts_response$groups[which(mut_counts_response$SNV_change_in_mutations<=v)]<-"Reduced_on_therapy"
mut_counts_response$groups[which(mut_counts_response$SNV_change_in_mutations>v)]<-"Increased_on_therapy"
mut_counts_response$groups<-as.factor(mut_counts_response$groups)
#table(mut_counts_response$groups)

OS_fit <- survfit(Surv(OS_time_Mon, OS_event) ~ groups, data = mut_counts_response)
#pdf(paste0(out_dir,"fig1F_OS.pdf"))
ggsurvplot(OS_fit, data = mut_counts_response,
           pval = TRUE, palette=c("red","blue"),
           risk.table = TRUE, xlab="Months",pval.coord=c(2,0.25),
           legend.labs=c("Increased_on_therapy", "Reduced_on_therapy"), legend="bottom",
           ylab="Overall Survival Probability",title="Survival split by median of SNV_change_in_mutations",
           font.title = c(8, "bold", "darkblue"))
#dev.off()

PFS_fit <- survfit(Surv(PFS_time, PFS_event) ~ groups, data = mut_counts_response)
#pdf(paste0(out_dir,"fig1G_PFS.pdf"))
ggsurvplot(PFS_fit, data = mut_counts_response,
           pval = TRUE,palette=c("red","blue"),
           risk.table = TRUE,xlab="Months",pval.coord=c(2,0.25),
           legend.labs=c("Increased_on_therapy", "Reduced_on_therapy"),legend="bottom",
           ylab="PFS",title="Survival split by median of SNV_change_in_mutations",
           font.title = c(8, "bold", "darkblue"))
#dev.off()

```

### Overall Survival for SNV and INDEL mut counts. 


```{r echo=FALSE, warning=FALSE,results='hide',include=FALSE}

comps<-c("SNV_pre_mutations","SNV_on_mutations","SNV_change_in_mutations","INDEL_pre_mutations", "INDEL_on_mutations","INDEL_change_in_mutations")

wfd2<-mut_counts_response
#table(wfd2$patient)
survcurvs<-lst()

for(i in comps){
val<-median(wfd2[,i])
wfd2$Highlow<-ifelse(wfd2[,i]>val, "high", "low")
wfd2$sHighlow<-as.factor(wfd2$Highlow)


f<-survfit(Surv(OS_time_Mon, OS_event) ~ Highlow, data = wfd2)

#pdf(paste(outdir, "survivalCurve", i, ".pdf", sep=''), height=6, width=6,onefile = FALSE )
survp<-ggsurvplot(fit = f,pval=TRUE, risk.table=TRUE,
                  xlab = "Months", 
                   legend.labs=c("High", "Low"),
                  legend="bottom",
                  palette=c("red", "blue"),
                  pval.coord=c(2,0.25),
                  ylab = "Overall survival probability", title=paste("Survival split by median of", i, sep=" "),
                  font.title = c(8, "bold", "darkblue"))
print(survp)
#dev.off()
survcurvs[i]<-survp

}

#print(c(survcurvs[1], survcurvs[2], survcurvs[3], survcurvs[4], survcurvs[5], survcurvs[6]))
```


### Figure 1 supplemental 
####	A: coOnco
####	B: coOnco
####	C: OddsRatio
####  D: OddsRatio
####  E: OncoPlot with response
####  F: survival based on HED